---
title: "Assignment 9--R"
author: "Insert your name here"
date: "September 5, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary libraries here.
```{r}
library(dplyr)
library(readr)
library(ggformula)
```

## Problem 1:  Colleges and Universities

In this problem, you will do further cleaning and analysis of the data from the 1995 US News and World Report on colleges and universities in the US.

a. In the Python portion of homework 9, you created a modified version of the data set usnews.csv.  Read the modified data into R.  Check the first few values of each vector to ensure that they were read correctly.

```{r}
usnews <- read_csv("usframe.csv")
head(usnews)
```

b. Examine the summary of each variable.  Identify any unrealistic values and set them to missing.

- It may be helpful to use control flow or functions to help organize your work.
```{r}
summary(usnews)
```

```{r}
which(usnews$`IQR:Math` < 0)
usnews$`IQR:Math`[usnews$`IQR:Math` == -10] <- NA
```

```{r}
usnews$`Student/faculty ratio`[usnews$`Student/faculty ratio` > 50] <- NA
```

```{r}

usnews$`Num applications received`[usnews$`Num applications received` > 40000] <- NA
usnews$`Num applicants accepted`[usnews$`Num applicants accepted` > 25000] <- NA
usnews$`Num students enrolled`[usnews$`Num students enrolled` == 4520] <- NA
```


```{r}
summary(usnews)
```


 **Write a sentence** describing what you did, naming the colleges or universities affected.  (For example, "Listed ages less than zero (ABC University, XYZ College) were converted to missing data.")

Pembroke State University had an IQR: Math value of -10 and was converted to missing data.

Northwood University has a student/faculty ration of 91.80 and was converted to missing data

St Leo College has a student/faculty ratio of 72.4 and was converted to missing data

Rutgers at New Brunswich received 48094 applications and accepted 26330 but only has 4520 students enrolled. Converted all three to missing data

c. Find the mean percentage of alumni who donate, for private and public schools.

```{r}
meanPub <- mean(usnews$`Pct alumni who donate`[usnews$Pub_or_Priv == "Public"], na.rm=T)
meanPub <- as.vector(meanPub)
```

```{r}
meanPriv <- mean(usnews$`Pct alumni who donate`[usnews$Pub_or_Priv == "Private"], na.rm=T)
meanPriv <- as.vector(meanPriv)
```

d. Test whether there is evidence that a higher percentage of alumni from private schools donate to their schools, compared to alumni from public schools.

- Hint:  In part c, we took the mean of this variable in each group.  What does this tell you about what type of hypothesis to use? two sample t-test

```{r}
t1 <- usnews$`Pct alumni who donate`[usnews$Pub_or_Priv == "Public"]
t1 <- t1[!is.na(t1)]
t2 <- usnews$`Pct alumni who donate`[usnews$Pub_or_Priv == "Private"]
t2 <- t2[!is.na(t2)]
```

```{r}
testResult <- t.test(t2, t1, alternative="greater")
testResult
```

**State your conclusion** in context.

H0: There's not higher percentage of donations from alumni between public and private schools

Ha: There is a higher percentage  of donations from alumni between private and public schools

Since the p-value < 2.2e-16 we can reject the null hypothesis and say that there is a higher percentage of alumni from private schools donating to their schools than public schools.

e. Use `write_csv()` or `write.csv()` to save your updated data set.

- If you are using `write.csv()`, consult the R documentation to set the arguments for the function.  Your output file should not have row names or row numbers.
- After you save the file, open it in Excel, Notepad++, or a program of your choice to verify that the version you saved contains the updates you made in part b.

```{r}
write.csv(usnews, "usnewsR.csv", row.names = FALSE)
```

## Problem 2:  Current Population Survey

The data set cps.csv contains data from the 1985 Current Population Survey.

- Dataset:  "Wages from the Current Population Survey," http://www.macalester.edu/~kaplan/ism/, from Daniel Kaplan, *Statistical Modeling: A Fresh Approach.* Original source:  Berndt, ER.  *The Practice of Econometrics 1991.*  Addison-Wesley.

- Metadata:  cps_metadata.pdf, from p. 418 of *Statistical Modeling: A Fresh Approach* by Daniel Kaplan.

a. Read the data into R and plot wages versus education.
```{r}
cps <- read_csv("cps.csv")
gf_point(wage ~ educ, data = cps)
```

**Comment** on the appropriateness of linear regression.
The wage vs education plot seems like it may have a linear relationship and we're looking at how a variable only compares to one other variable

b. Perform the linear regression and examine the diagnostic plots.
```{r}
cpsreg <- lm(wage ~ educ, data=cps)
summary(cpsreg)
```

```{r}
par(mfrow = c(2,2))
plot(cpsreg)
```

**Explain** why transforming the wages variable is a good idea in this case.

There's a lot of changes in variance in the Residuals vs Fitted plot so a transformation is a good idea.

c. The variable `wage` has units of dollars/hour.  Create a new variable, `time`, equal to 1/`wage`.  (So `time` has units of hours/dollar, or the length of time a person must work to earn $1.00.)

```{r}
cps$time <- 1/cps$wage
```

d. Plot time versus education.  
```{r}
gf_point(time ~ educ, data = cps)
```
Comment on the appropriateness of linear regression.

e. Perform the linear regression.
```{r}
timereg <- lm(time ~ educ, data = cps)
summary(timereg)
```

Based on these results, are you happy with your decision to pursue a master's degree?  **Explain.**

The regression makes it seem like the more education you have the more money you will end up making. I guess from that standpoint I'm glad I'm pursuing my master's even though its not really about the money for me.

f. Examine the diagnostic plots.
```{r}
par(mfrow = c(2,2))
plot(timereg)
```

**Which individuals** appear to be outliers on the residual vs. predicted plot?
216, 507, 518
Re-do the regression without those individuals.
```{r}
cps2 <- cps
cps2[216,] <- NA
cps2[507,] <- NA
cps2[518,] <- NA

timereg2 <- lm(time ~ educ, data = cps2)
summary(timereg2)
```

```{r}
par(mfrow = c(2,2))
plot(timereg2)
```

**Did excluding** the outliers change your conclusion?

No it did not. It still looks like the more education you have, the more money you'll end up making.

- Submit your .Rmd file, your knitted .docx or .pd f (.pdf preferred), and your updated data set from problem 1e to GitHub.